Missing Data: Current Practice in Football Research and Recommendations for Improvement

### 2020-10-30 paper/paper.Rmd
Nick orcid.

```{r load-funs}
pivot_var <- function(data, var){
  data %>% 
    select(title,
           reviewer,
           {{ var }}) %>% 
    pivot_wider(names_from = reviewer,
                values_from = {{ var }})
}

agree_var <- function(data, var){
  statement_agree <- pivot_var(data, {{ var }}) %>% 
    select(reviewer_1, reviewer_2) %>% 
    irr::agree()

  statement_agree$value
}

```{r read-data, message = FALSE, include = FALSE}
gs_football <- readr::read_rds(here::here("data",
                           "gs-football.rds"))
```

```{r mismatch, message = FALSE, include = FALSE}
# Now we bring the third and final review 
gs_mismatches <- readr::read_csv(here::here("data",
"missing-data-mismatches.csv"))

gs_mismatch_pivot <- gs_mismatches %>% 
  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
               names_to = "reviewer",
               values_to = "decision") %>% 
  pivot_wider(names_from = variable,
              values_from = decision)

gs_mismatch_pivot %>% 
  slice(1:3)

gs_football
```

```{r finalise-matches}
gs_football_final_matches <- gs_football %>% 
  pivot_longer(cols = any_missing_data_statement:was_data_shared,
               names_to = "variable",
               values_to = "value") %>% 
  pivot_wider(names_from = reviewer) %>% 
  left_join({ 
      select(gs_mismatches, 
             -reviewer_1,
             -reviewer_2)
    }, 
            by = c("variable", "title")) %>% 
  # double check that we get our data back
  # filter(reviewer_1 != reviewer_2)
  mutate(third_check = as.integer(third_check),
         final_check = case_when(
    is.na(third_check) ~ reviewer_1,
    reviewer_1 != reviewer_2 ~ third_check
  ))  %>%  select(-c(reviewer_1, 
            reviewer_2,
            third_check,
            notes)) %>% 
  pivot_wider(names_from = variable,
              values_from = final_check)
  # filter(reviewer_1 != reviewer_2)

```{r load-drake}
# loadd()
```

@@

ORCiD:

@@

Affiliations: 0000-0003-1460-8722

@@

```{r fun-pct-binom-conf}
percent_binom_conf <- function(binom_conf){
  mean_pct <- percent(binom_conf$mean, accuracy = 0.1)
lower_pct <- percent(binom_conf$lower, accuracy = 0.1)
upper_pct <- percent(binom_conf$upper, accuracy = 0.1)

return(data.frame(mean_pct = mean_pct,
                  lower_pct = lower_pct,
                  upper_pct = upper_pct))
}

```{r binom-ci-fun}
# need to update the gs_football
library(binom)

football_binom_confint <- function(data, var){

  var_vec <- data %>% pull({{ var }})
  successes <- sum(var_vec, na.rm = TRUE)
  trials <- sum(!is.na(var_vec))

  binom.confint(x = successes, 
              n = trials, 
              conf.level = 0.95, 
              methods = "exact")
}
```

@@

binom_ci_miss_reported <- football_binom_confint(gs_football_final_matches, 
                                                 any_missing_data_statement) %>% 
  percent_binom_conf()

@@

```{r afl-stats, cache = TRUE}

# Aim of this section:
# Simulate data from a real data set, insert missing
# 'Study aim': predicting hamstring strain injury, using number of sprints and sprint distance as x-variables

# Simulate dataset
# NOTE: fix this later in the drake plan
# d <- get_afltables_stats(start_date = "2020-08-01", end_date = "2020-09-01")
# write_rds(d, here::here("data", "afl-data.rds"))
d <- readr::read_rds(here::here("data", "afl-data.rds"))
afl_data <- d %>% select(ID, Round, Playing.for, Venue)

# Simulate injury and sprint variables
injury <- rbinom(n = nrow(d), 1, prob = 0.1)
sprints <- rpois(n = nrow(d), lambda = 100)
sprint_distance <- truncnorm::rtruncnorm(n = nrow(afl_data), 
                                         a = 0, 
                                         b = Inf, 
                                         mean = 2000, 
                                         sd = 500) # Check if this is reasonable (mean 2 km per game)

data <- cbind(injury, sprints, sprint_distance)

# Ampute missing values
data_ampute <- ampute(data = data[,1:3])
# try and tinker with the amount of missings - so the variables have different
# amounts of missingness
# try and introduce some missingness mechanisms

data_with_missing <- data_ampute$amp

# Put the dataset together
data_plot <- cbind(afl_data, data_with_missing)

# Clean variable names
data_plot <- data_plot %>% clean_names() %>% rename(team = playing_for)

```{r rename-afl-data}
data_plot_clean <- data_plot %>% 
  mutate(team = recode(team, 
                        "Gold Coast" = "GC",
                        "Greater Western Sydney" = "GWS",
                        "Fremantle" = "FRE",
                        "Collingwood" = "COL",
                        "Port Adelaide" = "PA",
                        "Western Bulldogs" = "WB",
                        "Richmond" = "RIC",
                        "Brisbane Lions" = "BL",
                        "Geelong" = "GEE",
                        "North Melbourne" = "NM",
                        "Adelaide" = "ADE",
                        "Melbourne" = "MEL",
                        "Sydney" = "SYD",
                        "St Kilda" = "STK",
                        "Essendon" = "ESS",
                        "West Coast" = "WC",
                        "Carlton" = "CAR",
                        "Hawthorn" = "HAW")) %>% 
  as_tibble() %>% 
  rename(sprint_dist = sprint_distance)

@@

aq_vis_miss <- vis_miss(data_plot_clean) + labs(tag = "(A)")
  #theme(text = element_text(size = 14))

aq_fct <- gg_miss_fct(data_plot_clean, team) + 
  #theme(text = element_text(size = 14)) +
  labs(x = "Team", y = "Variable", tag = "(B)") +
  theme(axis.text.x = element_text(size = 5))

gg_miss_scatter <- ggplot(data_plot_clean,
       aes(x = sprint_distance,
           y = sprints)) +
  geom_miss_point(size = 2) + 
  theme_minimal() +
  #theme(text = element_text(size = 14)) +
  labs(x = "Total sprint distance (m)", 
       y = "Number of sprints (count)", 
       colour = "",
       tag = "(C)")

aq_miss_var <- gg_miss_var(data_plot_clean) + labs(tag = "(D)")
```

>>>>

```{r load-targets}
loadd(gs_football)
loadd(gs_mismatches)
loadd(simulated_afl_data)
loadd(gs_football_final_matches)
```

```{r mismatch, message = FALSE, include = FALSE}
# Now we bring the third and final review 
gs_mismatch_pivot <- gs_mismatches %>% 
  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
               names_to = "reviewer",
               values_to = "decision") %>% 
  pivot_wider(names_from = variable,
              values_from = decision)
```

@@

ORCiD: 0000-0003-1460-8722

@@

Affiliations

@@

@@

binom_ci_miss_reported <- football_binom_confint(
  gs_football_final_matches, 
  any_missing_data_statement
  ) %>% 
  percent_binom_conf()

@@

@@

aq_vis_miss <- vis_miss(simulated_afl_data) + labs(tag = "(A)")
  #theme(text = element_text(size = 14))

aq_fct <- gg_miss_fct(simulated_afl_data, team) + 
  #theme(text = element_text(size = 14)) +
  labs(x = "Team", y = "Variable", tag = "(B)") +
  theme(axis.text.x = element_text(size = 5))

gg_miss_scatter <- ggplot(simulated_afl_data,
       aes(x = sprint_dist,
           y = sprints)) +
  geom_miss_point(size = 2) + 
  theme_minimal() +
  #theme(text = element_text(size = 14)) +
  labs(x = "Total sprint distance (m)", 
       y = "Number of sprints (count)", 
       colour = "",
       tag = "(C)")

aq_miss_var <- gg_miss_var(simulated_afl_data) + labs(tag = "(D)")
```
